% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here




\begin{document}


\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf DATA 621 - Final Project}

  \author{
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf DATA 621 - Final Project}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Use 250 words or less to summarize your problem, methodology, and major
outcomes. \newline This is the second line of abstract.
\end{abstract}

\noindent%
{\it Keywords:} keywordA, keywordB
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Describe the background and motivation of your problem.

\hypertarget{literature-review}{%
\subsection{Literature Review}\label{literature-review}}

Discuss how other researchers have addressed similar problems, what
their achievements are, and what the advantage and drawbacks of each
reviewed approach are. Explain how your investigation is similar or
different to the state-of-theart. Please cite the relevant papers where
appropriate.

\hypertarget{methodology}{%
\subsection{Methodology}\label{methodology}}

Discuss the key aspects of your problem, data set and regression
model(s). Given that you are working on real-world data, explain at a
high-level your exploratory data analysis, how you prepared the data for
regression modeling, your process for building regression models, and
your model selection.

\hypertarget{experimentation-and-results}{%
\subsection{Experimentation and
Results}\label{experimentation-and-results}}

Describe the specifics of what you did (data exploration, data
preparation, model building, model selection, model evaluation, etc.),
and what you found out (statistical analyses, interpretation and
discussion of the results, etc.).

\hypertarget{discussion-and-conclusions}{%
\subsection{Discussion and
Conclusions}\label{discussion-and-conclusions}}

Conclude your findings, limitations, and suggest areas for future work.

\hypertarget{references}{%
\subsection{References}\label{references}}

Be sure to cite all references used in the report (APA format).

\hypertarget{appendices}{%
\subsection{Appendices}\label{appendices}}

\begin{itemize}
\tightlist
\item
  Supplemental tables and/or figures.
\item
  R statistical programming code.
\end{itemize}

\bibliographystyle{agsm}
\bibliography{}

\end{document}
