---
title: "A Logistic Regression Approach to CoIL Challenge 2000"
author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
bibliography: "Manuscript.bib"
output: 
  word_document:
    reference_docx: word-template.docx
---

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(pander)

prepare_data <- function(df){
  names(df) <- c(
    "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
    "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN",
    "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG",
    "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA",
    "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2",
    "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
    "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", 
    "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT",  "PAANHANG",
    "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG",
    "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED",
    "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
    "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM",
    "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
    "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")
  
  MOSTYPE_labels <- c(
    "1" = "1 High Income, expensive child",
    "2" = "2 Very Important Provincials",
    "3" = "3 High status seniors",
    "4" = "4 Affluent senior apartments",
    "5" = "5 Mixed seniors",
    "6" = "6 Career and childcare",
    "7" = "7 Dinki's (double income no kids)",
    "8" = "8 Middle class families",
    "9" = "9 Modern, complete families",
    "10" = "10 Stable family",
    "11" = "11 Family starters",
    "12" = "12 Affluent young families",
    "13" = "13 Young all american family",
    "14" = "14 Junior cosmopolitan",
    "15" = "15 Senior cosmopolitans",
    "16" = "16 Students in apartments",
    "17" = "17 Fresh masters in the city",
    "18" = "18 Single youth",
    "19" = "19 Suburban youth",
    "20" = "20 Ethnically diverse",
    "21" = "21 Young urban have-nots",
    "22" = "22 Mixed apartment dwellers",
    "23" = "23 Young and rising",
    "24" = "24 Young, low educated",
    "25" = "25 Young seniors in the city",
    "26" = "26 Own home elderly",
    "27" = "27 Seniors in apartments",
    "28" = "28 Residential elderly",
    "29" = "29 Porchless seniors: no front yard",
    "30" = "30 Religious elderly singles",
    "31" = "31 Low income catholics",
    "32" = "32 Mixed seniors",
    "33" = "33 Lower class large families",
    "34" = "34 Large family, employed child",
    "35" = "35 Village families",
    "36" = "36 Couples with teens 'Married with children'",
    "37" = "37 Mixed small town dwellers",
    "38" = "38 Traditional families",
    "39" = "39 Large religous families",
    "40" = "40 Large family farms",
    "41" = "41 Mixed rurals")
  
  MGEMLEEF_labels <- c(
    "1" = "20-30 years",
    "2" = "30-40 years",
    "3" = "40-50 years",
    "4" = "50-60 years",
    "5" = "60-70 years",
    "6" = "70-80 years")
  
  MOSHOOFD_labels <- c(
    "1" = "Successful hedonists",
    "2" = "Driven Growers",
    "3" = "Average Family",
    "4" = "Career Loners",
    "5" = "Living well",
    "6" = "Cruising Seniors",
    "7" = "Retired and Religeous",
    "8" = "Family with grown ups",
    "9" = "Conservative families",
    "10" = "Farmers")
  
  MGODRK_labels <- c(
    "0" = "0%",
    "1" = "1 - 10%",
    "2" = "11 - 23%",
    "3" = "24 - 36%",
    "4" = "37 - 49%",
    "5" = "50 - 62%",
    "6" = "63 - 75%",
    "7" = "76 - 88%",
    "8" = "89 - 99%",
    "9" = "100%")
  
  PWAPART_labels <- c(
    "0" = "f 0",
    "1" = "f 1 - 49",
    "2" = "f 50 - 99",
    "3" = "f 100 - 199",
    "4" = "f 200 - 499",
    "5" = "f 500 - 999",
    "6" = "f 1000 - 4999",
    "7" = "f 5000 - 9999",
    "8" = "f 10,000 - 19,999",
    "9" = "f 20,000 - ?")
  
  set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)
  
  df %>%
    mutate(CUSTOMER_TYPE = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
    mutate(CUSTOMER_TYPE = as.factor(CUSTOMER_TYPE)) %>%
    mutate(MOSTYPE = as.factor(MOSTYPE),
           MGEMLEEF = as.factor(MGEMLEEF),
           MOSHOOFD = as.factor(MOSHOOFD),
           MGODRK = as.factor(MGODRK),
           PWAPART = as.factor(PWAPART),
           CARAVAN = as.factor(CARAVAN)) %>%
    mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
           MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
           MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
           MGODRK = recode(MGODRK, !!!MGODRK_labels),
           PWAPART = recode(PWAPART, !!!PWAPART_labels))
}

eval <- read.delim("data/ticeval2000.txt", header = FALSE)
temp <- read.delim("data/tictgts2000.txt", header = FALSE)
eval$CARAVAN <- temp$V1
eval <- prepare_data(eval)
df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
```

# Abstract

A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics. 

*Key words: CoIL Challenge, Logistic Regression*

# Introduction

Businesses use data science to extract insights from data.  One pratical application is identifying households to include in a marketing campaign.  In this paper we set out to identify potential customers for an insurance product using real world data from the Computational Intelligence and Learning (CoIL) Challenge. 

# Literature Review

<!--Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-the-art. Please cite the relevant papers where appropriate.-->


Fourty-three other research teams have attempted to identify potential caravan insurance customers [@Putten].  They used a variety of approaches including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity", fill = "#377eb8") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122) +
  scale_fill_brewer(palette = "Set1")
```

The maximum number of potential policy owners that could be found is 238.  Previous researchers identified 95 policy owners on average.  The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.  A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].

# Methodology

<!-- Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection. -->

# Appendix

## Data Dictionary

```{r, results='asis', echo=FALSE}
data_dictionary <- data.frame(
  Name = c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
           "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV",
           "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD",
           "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD",
           "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD",
           "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART",
           "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM",
           "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT",
           "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
           "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
           "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR",
           "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG",
           "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG",
           "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
           "ABYSTAND", "CARAVAN"),
  Description = c("Customer Subtype", "Number of houses", "Avg size household",
                  "Avg age", "Customer main type", "Roman catholic",
                  "Protestant", "Other religion", "No religion", "Married",
                  "Living together", "Other relation", "Singles",
                  "Household without children", "Household with children",
                  "High level education", "Medium level education",
                  "Lower level education", "High status", "Entrepreneur",
                  "Farmer", "Middle management", "Skilled labourers",
                  "Unskilled labourers", "Social class A", "Social class B1",
                  "Social class B2", "Social class C", "Social class D",
                  "Rented house", "Home owners", "1 car", "2 cars", "No car",
                  "National Health Service", "Private health insurance",
                  "Income < 30.000", "Income 30-45.000", "Income 45-75.000",
                  "Income 75-122.000", "Income >123.000", "Average income",
                  "Purchasing power class",
                  "Contribution private third party insurance",
                  "Contribution third party insurance (firms)",
                  "Contribution third party insurane (agriculture)",
                  "Contribution car policies",
                  "Contribution delivery van policies",
                  "Contribution motorcycle/scooter policies",
                  "Contribution lorry policies",
                  "Contribution trailer policies",
                  "Contribution tractor policies",
                  "Contribution agricultural machines policies",
                  "Contribution moped policies",
                  "Contribution life insurances",
                  "Contribution private accident insurance policies",
                  "Contribution family accidents insurance policies",
                  "Contribution disability insurance policies",
                  "Contribution fire policies",
                  "Contribution surfboard policies",
                  "Contribution boat policies",
                  "Contribution bicycle policies",
                  "Contribution property insurance policies",
                  "Contribution social security insurance policies",
                  "Number of private third party insurance",
                  "Number of third party insurance (firms)",
                  "Number of third party insurane (agriculture)",
                  "Number of car policies",
                  "Number of delivery van policies",
                  "Number of motorcycle/scooter policies",
                  "Number of lorry policies",
                  "Number of trailer policies", "Number of tractor policies",
                  "Number of agricultural machines policies",
                  "Number of moped policies", "Number of life insurances",
                  "Number of private accident insurance policies",
                  "Number of family accidents insurance policies",
                  "Number of disability insurance policies",
                  "Number of fire policies", "Number of surfboard policies",
                  "Number of boat policies", "Number of bicycle policies",
                  "Number of property insurance policies",
                  "Number of social security insurance policies",
                  "Number of mobile home policy"))
pander(data_dictionary)
```

\newpage

## R statistical programming code.

<code>
```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
conn <- file("CoIL.r", open = "r")
lines <- readLines(conn)
n_lines <- length(lines)
for (i in 1:n_lines) {
  cat(lines[i], "\n")
}
close(conn)
```
</code>

## References