---
title: "Exploratory Data Analysis"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(gridExtra)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r read_data}
training <- read.csv("../data/moneyball-training-data.csv")
```



## Exploration

### Summary Statistics

The following is a summary of the data.

```{r}
training %>% 
  select(-INDEX) %>%
  summary(.) %>%
  kable() %>%
  kable_styling()
```

### Plots

The following density plots show the spread of the data.  The red verticle line is the mean and the blue verticle line is the median.  The scatter plot shows the relationship between wins and the variable

```{r plot_me}
plot_me <- function(d, title, xlab, ylab){
  nas <- d %>% filter(is.na(x)) %>% nrow()
  density <- d %>%
    filter(!is.na(x)) %>%
    ggplot(., aes(x, stat(count))) +
    geom_density() +
    labs(title = title, 
       x = xlab,
       y = ylab) +
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray95")) +
    geom_vline(xintercept = mean(d$x, na.rm = TRUE), color = "red") +
    geom_vline(xintercept = median(d$x, na.rm = TRUE), color = "blue")
  # Add annotation for NA's
  if(nas > 0){
    density <- density + ggtitle(title, subtitle = paste(nas, "Null Observations")) 
  }
  
  # Create a scatter plot
  scatter <- ggplot(d, aes(x, TARGET_WINS)) +
    geom_point() + 
    geom_smooth(method = lm) + 
    labs(x = xlab,
         y = ylab) +
    expand_limits(y=0) +
    theme(panel.background = element_blank(),
          panel.grid.major = element_line(color = "gray95"))
  # Put them side by side
  grid.arrange(density, scatter, ncol=2)
}
```

```{r}
training %>%
  mutate(x = TARGET_WINS) %>%
  plot_me(., "Wins", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_H) %>%
  plot_me(., "Base Hits (TEAM_BATTING_H)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_2B) %>%
  plot_me(., "Doubles (TEAM_BATTING_2B)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_3B) %>%
  plot_me(., "Triples (TEAM_BATTING_3B)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_HR) %>%
  plot_me(., "Homeruns (TEAM_BATTING_HR)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_BB) %>%
  plot_me(., "Walks (TEAM_BATTING_BB)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_HBP) %>%
  plot_me(., "Hit by Pitch (TEAM_BATTING_HBP)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BATTING_SO) %>%
  plot_me(., "Strikeouts (TEAM_BATTING_SO)", element_blank(), element_blank())
```


```{r}
training %>%
  mutate(x = TEAM_BASERUN_SB) %>%
  plot_me(., "Stolen Bases (TEAM_BASERUN_SB)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_BASERUN_CS) %>%
  plot_me(., "Caught Stealing Bases (TEAM_BASERUN_CS)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_FIELDING_E) %>%
  plot_me(., "Errors (TEAM_FIELDING_E)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_FIELDING_DP) %>%
  plot_me(., "Double Plays (TEAM_FIELDING_DP)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_BB) %>%
  plot_me(., "Walks Allowed (TEAM_PITCHING_BB)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_H) %>%
  plot_me(., "Hits Allowed (TEAM_PITCHING_H)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_HR) %>%
  plot_me(., "Homeruns Allowed (TEAM_PITCHING_HR)", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING_SO) %>%
  plot_me(., "Strikouts by Pitcher (TEAM_PITCHING_SO)", element_blank(), element_blank())
```

## Missing Data

### Batting Strike Outs

To fill the missing in the `r nrow(training[is.na(training$TEAM_BATTING_SO),])` missing data we will alternate between the two modes (578 and 909)

```{r}
## How did we know that? We looked at this table
# D <- training[!is.na(training$TEAM_BATTING_SO),]$TEAM_BATTING_SO
# data.frame(x = density(D)$x, y = density(D)$y) %>%
#   arrange(desc(y)) %>%
#   kable() %>%
#   kable_styling()
# # Modes at
# # 578 and 909
```

```{r}
set.seed(42)

training <- training %>%
  mutate(TEAM_BATTING_SO_flag  = ifelse(is.na(TEAM_BATTING_SO), 1, 0)) %>%
  mutate(TEAM_BATTING_SO  = ifelse(is.na(TEAM_BATTING_SO), sample(c(578, 909), 1), TEAM_BATTING_SO))
```

### Pitching Strike Outs

To fill the the `r nrow(training[is.na(training$TEAM_PITCHING_SO),])` missing values with the mean.

```{r}
training <- training %>%
  mutate(TEAM_PITCHING_SO_flag = ifelse(is.na(TEAM_PITCHING_SO), 1, 0)) %>%
  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_PITCHING_SO), mean(TEAM_PITCHING_SO), TEAM_PITCHING_SO))
```

### Double Plays

To fill the the `r nrow(training[is.na(training$TEAM_FIELDING_DP),])` missing values with the mean.

```{r}
training <- training %>%
  mutate(TEAM_FIELDING_DP_flag = ifelse(is.na(TEAM_FIELDING_DP), 1, 0)) %>%
  mutate(TEAM_PITCHING_SO = ifelse(is.na(TEAM_FIELDING_DP), mean(TEAM_FIELDING_DP), TEAM_FIELDING_DP))
```

## Scaled and Combined

The idea behind this model is that teams that are better than the average will win more games and teams worse than the average will win less.  The way we determine if a team is better than average is by looking at how well they preform at batting, pitching, and fielding.

Since there are more than one way to win a baseball game (i.e. have some power sluggers that hit home runs, vs have really good single batters.) we need to combine the various batting measures.  Now since getting a strikout at bat is bad, we need to change the sign of this variable.  That way it can be combined and will fit the better teams win more and worse teams less model.

We are going to scale all variables .  That centers them at 0 and gives them a standard deviation of 1.  We can then combine almost all the batting variables into one measure (hit by pitcher is excluded).

```{r}
training <- training %>%
  mutate(TEAM_BATTING_SO = -1 * TEAM_BATTING_SO) %>%
  mutate(TEAM_PITCHING_SO = -1 * TEAM_PITCHING_SO) %>%
  mutate(TEAM_PITCHING_H = -1 * TEAM_PITCHING_H) %>%
  mutate(TEAM_FIELDING_E = -1 * TEAM_FIELDING_E) %>%
  mutate_at(scale, .vars = vars(-TARGET_WINS)) %>%
  #mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
  mutate(TEAM_BATTING = TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_SO) %>%
  mutate(TEAM_PITCHING = TEAM_PITCHING_BB + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) %>%
  mutate(TEAM_FIELDING = TEAM_FIELDING_DP + TEAM_FIELDING_E)
```

```{r}
training %>%
  mutate(x = TEAM_BATTING) %>%
  plot_me(., "Batting", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_PITCHING) %>%
  plot_me(., "Pitching", element_blank(), element_blank())
```

```{r}
training %>%
  mutate(x = TEAM_FIELDING) %>%
  plot_me(., "Fielding", element_blank(), element_blank())
```

```{r}
fit <- lm(TARGET_WINS ~ TEAM_BATTING + TEAM_PITCHING + TEAM_FIELDING, training)
summary(fit)
```

This model says that the average baseball team will win about `r round(coef(fit)[1],0)` games.  If their batting is one standard deviation better than the average they will win `r round(coef(fit)[2],0)` more games.  They will win `r round(coef(fit)[3],0)` more games for if their pitching and `r round(coef(fit)[4],0)` if their fielding is better than average.